Information Transfer in Linear Multi-Task Learning: High-dimensional Asymptotics and Implications

### 2020-09-03 arxiv/transfer.tex
wip
edits

\section{Illustrative Examples for Isotropic Features}\label{sec_special}

%We provide precise explanations to the phenomena of negative transfer in multi-task learning.
%We provide tight bounds on the bias and variance of the multi-task estimator for two tasks.
The results in Section \ref{sec_general} show when multi-task learning outperforms single-task learning for general task covariance matrices and ground truth model parameters.
In particular, the general conditions depend on specific properties of task data such as the covariate shift matrix.
However, the conditions given by the limiting bias and variance of multi-task learning can be difficult to interpret.
The goal of this section is to address this issue.
We interpret the limiting bias and variance equations in an isotropic covariance setting that is described below and a simplified covariate-shifted setting.
We show that three properties of task data affect the performance of multi-task learning as follows.
(i) \textit{Task similarity}: We explain the phenomenon of negative transfer precisely as tasks become more different.
(ii) \textit{Sample ratio}: We further explain a curious phenomenon where increasing the sample ratio between the source and target task helps initially, but hurts eventually.
(iii) \textit{Covariate shift}: Finally, we show that as the sample size of the source task increases, we show that the covariate shift worsens the performance of the multi-task estimator.
%Finally, we provide a extend our results from two tasks to many tasks with the same features.
%We explain from three perspectives, including \textit{task similarity}, \textit{sample size} and \textit{covariate shift}.
%We show how negative transfer occurs by varying task similarity or sample size.
%Then we show that when source task sample size becomes large, covariate shift causes more negative effects.

\smallskip
\noindent\textbf{The isotropic covariance setting.}
	For the two-task case, suppose that the covariates of both tasks have isotropic covariance matrices, that is, $\Sigma_1 = \Sigma_2 = \id$.
%	Each task has sample size $n_1 = \rho_1 \cdot p$ and $n_2 = \rho_2 \cdot p$.
	%And $X_1\in\real^{n_1\times p}, X_2\in\real^{n_2\times p}$ denote the covariates of the two tasks, respectively.
	Assume that for task two, its model parameter, denoted by $\beta_2$, consist of i.i.d. entries with mean zero and variance $\kappa^2$.
	For task one, its model parameter, denoted by $\beta_1$, is equal to $\beta_2$ plus i.i.d. entries with mean $0$ and variance $d^2$.
%	The labels are $Y_i = X_i\beta_i + \varepsilon_i$, where $\e_i$ consists of i.i.d. entries with mean zero and variance $\sigma^2$.
%	For our purpose, it is enough to think of the order of $d$ being $1/\sqrt{p}$ and $pd^2/\sigma^2$ being constant.
%The more precise conditions on the relations between $d^2$, $\sigma^2$ and $\kappa^2$ are given in  \eqref{choiceofpara}.
%	We assume that all the random variables have subexponential decay, while keeping in mind that our results can be applied under weaker moments assumptions as shown in Appendix \ref{sec_maintools}.

>>>>

\section{Illustrative Examples in an Isotropic Setting}\label{sec_special}

%We provide precise explanations to the phenomena of negative transfer in multi-task learning.
%We provide tight bounds on the bias and variance of the multi-task estimator for two tasks.
%The results in Section \ref{sec_general} show when multi-task learning outperforms single-task learning for general task covariance matrices and ground truth model parameters.
%In particular, the general conditions depend on specific properties of task data such as the covariate shift matrix.
%However, the conditions given by the limiting bias and variance of multi-task learning can be difficult to interpret.
%The goal of this section is to address this issue.
%We interpret the limiting bias and variance equations in an isotropic covariance setting that is described below and a simplified covariate-shifted setting.
%We show that three properties of task data affect the performance of multi-task learning as follows.
%(i) \textit{Task similarity}: We explain the phenomenon of negative transfer precisely as tasks become more different.
%(ii) \textit{Sample ratio}: We further explain a curious phenomenon where increasing the sample ratio between the source and target task helps initially, but hurts eventually.
%(iii) \textit{Covariate shift}: Finally, we show that as the sample size of the source task increases, we show that the covariate shift worsens the performance of the multi-task estimator.
%Finally, we provide a extend our results from two tasks to many tasks with the same features.
%We explain from three perspectives, including \textit{task similarity}, \textit{sample size} and \textit{covariate shift}.
%We show how negative transfer occurs by varying task similarity or sample size.
%Then we show that when source task sample size becomes large, covariate shift causes more negative effects.
We illustrate our main results (to be described in Section \ref{sec_general}) in a simplified setting.
For this setting, we simulate several empirical phenomena that are well-known in hard parameter sharing approaches.
We use our results to explain the empirical phenomena.

\medskip
\noindent\textbf{Isotropic features and models.}
	Imagine that we have two tasks and the covariates of both tasks have isotropic covariance matrices.
%	Each task has sample size $n_1 = \rho_1 \cdot p$ and $n_2 = \rho_2 \cdot p$.
	%And $X_1\in\real^{n_1\times p}, X_2\in\real^{n_2\times p}$ denote the covariates of the two tasks, respectively.
	For the ground truth linear model, we consider a random-effects model where the two tasks differ by an independent Gaussian vector.
	Specially, for task two, we assume that its model parameter, denoted by $\beta_2$, consists of i.i.d. Gaussian entries with mean zero and variance $\kappa^2$.
	For task one, we assume that its model parameter, denoted by $\beta_1$, is equal to $\beta_2$ plus i.i.d. Gaussian entries with mean $0$ and variance $d^2$.
	Therefore, the two tasks have different ground truth models. In particular, the distance between $\beta_1$ and $\beta_2$ roughly scales as $d^2 \cdot p$.

Using our techniques described later in Section \ref{sec_general}, we derive the limiting prediction loss of the hard parameter sharing estimator.
Without loss of generality, we describe the result for task two and remark that a similar result for task one holds.

\begin{corollary}[Derived from Lemma \ref{lem_cov_shift} and Lemma \ref{lem_cov_derivative}]
	In the isotropic setting described above, the prediction loss of the hard parameter sharing estimator for task two satisfies:
	\[ L(\hat{\beta}_2^{\MTL}) = ?? \]
\end{corollary}

%	The labels are $Y_i = X_i\beta_i + \varepsilon_i$, where $\e_i$ consists of i.i.d. entries with mean zero and variance $\sigma^2$.
%	For our purpose, it is enough to think of the order of $d$ being $1/\sqrt{p}$ and $pd^2/\sigma^2$ being constant.
%The more precise conditions on the relations between $d^2$, $\sigma^2$ and $\kappa^2$ are given in  \eqref{choiceofpara}.
%	We assume that all the random variables have subexponential decay, while keeping in mind that our results can be applied under weaker moments assumptions as shown in Appendix \ref{sec_maintools}.
